Manage data to get a number on prescribed burns in the Western US
Tyler L. McIntosh
CU Boulder CIRES Earth Lab
Last updated: 11/15/23

This script uses the following naming conventions wherever possible:
 lowerCamelCase for variables
 period.separated for functions
 underscore_separated for files

# Setup

```{r, message = FALSE, warning = FALSE, include = FALSE}
## Libraries ----

#Check the required libraries and download if needed
list.of.packages <- c("tidyverse", #Includes ggplot2, dplyr, tidyr, readr, purrr, tibble, stringr, forcats
                      "terra",
                      "sf",
                      "mapview",
                      "here",
                      "future", "future.apply", "furrr", "doFuture", "progressr", #Futureverse!
                      "tictoc", 
                      "mblm", #Median-based linear models (i.e. thiel-sen)
                      "plyr",
                      "scales", #add commas to ggplot axis
                      "tigris", #US data
                      "scales") #add commas to axis

#Install all packages that aren't installed yet
install.packages(setdiff(list.of.packages, rownames(installed.packages())))

#Load all packages
invisible(lapply(list.of.packages, library, character.only = TRUE)) #use 'invisible' to prevent output being displayed

## Clean workspace & set up environment ----
rm(list=ls()) #Ensure empty workspace if running from beginning
here::i_am("code/02_a_number_on_prescribed_burns_data.Rmd")
here::here() #Check here location
#OR
#setwd() #Set working directory directly
options(digits = 6) #Set standard decimal print output
options(scipen = 999) #Turn scientific notation on and off (0 = on, 999 = off)

#Ensure necessary directories are created
derivedDatDir <- here::here("data", "derived")
if (!dir.exists(derivedDatDir)){
  dir.create(derivedDatDir)
}

```

# Import US data

```{r}

# #Get shapefiles for the US
# # Fetch all US states
# usa <- tigris::states() %>% 
#   st_transform(st_crs(4269))
# 
# # Filter for contiguous states (excluding Alaska and Hawaii)
# conus <- usa[usa$STUSPS %in% c("WA", "ID", "MT", "ND", "MN", "MI", "ME", "WI", "VT", "NH", "NY", "SD", "WY", "OR", "IA", "NE", "IL", "NV", "UT", "CO", "KS", "OK", "NM", "AZ", "AR", "MO", "MS", "AL", "GA", "SC", "NC", "TN", "KY", "IN", "OH", "WV", "VA", "PA", "MD", "DE", "NJ", "CT", "RI", "MA", "CT", "CA", "TX", "FL", "LA"),]


westAbbrList <- c("WA", "OR", "CA", "ID", "MT", "WY", "NV", "AZ", "CO", "NM", "UT")
#west <- usa[usa$STUSPS %in% westAbbrList,]


```

# Import prescribed fire datasets (two different ones)

```{r}
#load nfpors
nfpors <- sf::st_read(here::here('data', 'raw', 'NFPORS_WestStates_2010_2021', 'NFPORS_WestStates_2010_2021.gdb'),
                      layer = "West_NFPORS_2010_2021")|>
  dplyr::filter(!is.na(ACTUALCOMPLETIONDATE)) #ensure all included burns were actually done

#load facts
facts <- sf::st_read(here::here('data', 'raw', 'S_USA.Activity_HazFuelTrt_PL.gdb', 'S_USA.Activity_HazFuelTrt_PL.gdb'),
                      layer = "Activity_HazFuelTrt_PL")

```

# Explore nfpors

```{r, eval = FALSE}

head(nfpors)
unique(nfpors$KEYPOINTNAME)
unique(nfpors$STATENAME) #state filter has been done correctly
unique(nfpors$TYPENAME) #includes Machine Pile Burn, Broadcast Burn, Fire Use, Jackpot Burn, and Hand Pile Burn
unique(nfpors$UNITOFMEAS)
unique(nfpors$WUIID)

#mapview(nfpors) #this is point data!

```


# Manipulate NFPORS data

```{r}

unique(nfpors$actualcompletionyear)
unique(year(as.Date(nfpors$ACTUALCOMPLETIONDATE)))
#actualcompletionyear has a few errors; make new one from the good data
nfpors <- nfpors %>%
  dplyr::mutate(ACTUALCOMPLETIONYEARNEW = year(as.Date(ACTUALCOMPLETIONDATE)))


```
# Explore facts, filter, manipulate

```{r, include = FALSE}
head(facts)
unique(facts$METHOD)
unique(facts$TREATMENT_TYPE)

yearsOfInterest = seq(2010,2021)

#Get only prescribed burning activities in the western US in the years we have NFPORS data
westPBFacts <- facts %>%
  dplyr::mutate(ACTUALCOMPLETIONYEARNEW = year(as.Date(DATE_COMPLETED))) |>
  dplyr::filter(STATE_ABBR %in% westAbbrList) |>
  dplyr::filter(TREATMENT_TYPE %in% c("Machine Pile Burn", "Jackpot Burn", "Fire Use", "Broadcast Burn", "Hand Pile Burn")) |>
  dplyr::filter(ACTUALCOMPLETIONYEARNEW %in% yearsOfInterest)

unique(westPBFacts$ACTUALCOMPLETIONYEARNEW)

```

# Prepare to compare NFPORS FACTS & raw FACTS

```{r}

nfporsAllSummary <- nfpors |>
  dplyr::group_by(STATEABBR, ACTUALCOMPLETIONYEARNEW) |>
  dplyr::summarize(NFPORS_all = sum(TOTALACCOMPLISHMENT)) |>
  sf::st_drop_geometry()

#filter NFPORS to just facts
nfporsFACTS <- nfpors |> dplyr::filter(SOURCE == "FACTS")

nfporsFACTSSummary <- nfporsFACTS |>
  dplyr::group_by(STATEABBR, ACTUALCOMPLETIONYEARNEW) |>
  dplyr::summarize(NFPORS_FACTS = sum(TOTALACCOMPLISHMENT)) |>
  sf::st_drop_geometry()

factsSummary <- westPBFacts |>
  as.data.frame() |>
  dplyr::group_by(STATE_ABBR, ACTUALCOMPLETIONYEARNEW) |>
  dplyr::summarize(FACTS_gis = sum(GIS_ACRES, na.rm=TRUE),
            FACTS = sum(NBR_UNITS_ACCOMPLISHED))

#Get summary of data
allSummaryRaw <- left_join(nfporsFACTSSummary, factsSummary,
                           by = c('STATEABBR' = 'STATE_ABBR', 'ACTUALCOMPLETIONYEARNEW' = 'ACTUALCOMPLETIONYEARNEW')) %>%
  left_join(nfporsAllSummary,
            by = c('STATEABBR' = 'STATEABBR', 'ACTUALCOMPLETIONYEARNEW' = 'ACTUALCOMPLETIONYEARNEW')) |>
  dplyr::mutate(FACTSGIS_NONGIS_DIFF = FACTS - FACTS_gis,
         FACTS_NFPORS_DIFF = FACTS - NFPORS_FACTS)

#Pivot longer
allSummaryRawLong <- allSummaryRaw %>% 
  dplyr::select(-FACTSGIS_NONGIS_DIFF, -FACTS_NFPORS_DIFF) |>
  tidyr::pivot_longer(cols = c(FACTS_gis,
                               FACTS,
                               NFPORS_FACTS,
                               NFPORS_all),
                                                    names_to = "datNm",
                                                    values_to = "acres") |>
  dplyr::mutate(hectares = acres * 0.404686) #convert to hectares

#write out dataset
write_csv(allSummaryRawLong, here::here(derivedDatDir, "rx_dataset_comparison_long_version.csv"))



```

# Create NFPORS + LCMS data

LCMS & FRG data pulled from GEE:
https://code.earthengine.google.com/60989ec067bec112f59c9f97d8f86a88

Read in data and reshape, merge with raw NFPORS

```{r}
#Read in data from GEE
geeNfporsDats <- read_csv(here::here('data', 'derived', 'gee_nfpors.csv')) |>
  dplyr::select(-`.geo`, -`system:index`) |>
  dplyr::mutate(name = paste0("LC_", LandCover_Year))

#LCMS classes; for context
lcmsClasses <- cbind(
  seq(1,15),
  c("Trees",
    "Tall Shrubs & Trees Mix (SEAK Only)",
    "Shrubs & Trees Mix",
    "Grass/Forb/Herb & Trees Mix",
    "Barren & Trees Mix",
    "Tall Shrubs (SEAK Only)",
    "Shrubs",
    "Grass/Forb/Herb & Shrubs Mix",
    "Barren & Shrubs Mix",
    "Grass/Forb/Herb",
    "Barren & Grass/Forb/Herb Mix",
    "Barren or Impervious",
    "Snow or Ice",
    "Water",
    "Non-Processing Area Mask")
) |>
  as.data.frame() |> 
  `names<-`(c("LandCover", "LandCoverDescription")) |>
   dplyr::mutate(LandCover = as.double(LandCover))

#FRG classes; for context
frgClasses <- cbind(
  seq(1,3),
  c("frcLowMix",
    "frcReplace",
    "frcOther")) |>
  as.data.frame() |>
  `names<-`(c("FRG", "FRGDescription"))  |>
   dplyr::mutate(FRG = as.double(FRG))

#Pivot wider
shortGeeNfporsDats <- geeNfporsDats |>
  dplyr::select(-LandCover_Year) |>
  tidyr::pivot_wider(names_from = name, values_from = LandCover)

#Join together and add column with land cover in burn year & landcover name, as well as FRG full name
nfporsWithGEE <- nfpors|>
  dplyr::mutate(PointId = objectid__) |>
  dplyr::left_join(shortGeeNfporsDats, by = c("PointId")) |>
  dplyr::mutate(index = paste0("LC_", ACTUALCOMPLETIONYEARNEW)) |> #THIS AND NEXT TWO LINES DO ROWWISE COMPUTE FOR THE LC_BURNYEAR
  dplyr::rowwise() |>
  dplyr::mutate(LC_BurnYear = get(index)) |>
  dplyr::ungroup() |>
  as.data.frame() |>
  dplyr::left_join(lcmsClasses, by = c("LC_BurnYear" = "LandCover")) |> #join landcover descriptions
  dplyr::left_join(frgClasses, by = c("FRG")) |> #join FRG descriptions
  dplyr::mutate(TOTALACCOMPLISHMENT_HA = TOTALACCOMPLISHMENT * 0.404686) #convert new HA column
  
#write out dataset
write_csv(nfporsWithGEE, here::here(derivedDatDir, "nfpors_with_gee.csv"))


```


