Create spatial graphics for good fire
Tyler L. McIntosh
CU Boulder CIRES Earth Lab
Last updated: 12/12/23

This script uses the following naming conventions wherever possible:
 lowerCamelCase for variables
 period.separated for functions
 underscore_separated for files

# Setup

```{r, message = FALSE, warning = FALSE, include = FALSE}
## Libraries ----

#Check the required libraries and download if needed
list.of.packages <- c("tidyverse", #Includes ggplot2, dplyr, tidyr, readr, purrr, tibble, stringr, forcats
                      "terra",
                      "sf",
                      "mapview",
                      "here",
                      "future", "future.apply", "furrr", "doFuture", "progressr", #Futureverse!
                      "tictoc", 
                      "mblm", #Median-based linear models (i.e. thiel-sen)
                      "plyr",
                      "kableExtra", #nice tables
                      "googlesheets4", #read in google sheet data
                      "scales", #add commas to ggplot axis
                      "tigris", #US data
                      "tmap", #mapping
                      "gridExtra", #multi-frame plots
                      "scales") #add commas to axis

#Install all packages that aren't installed yet
install.packages(setdiff(list.of.packages, rownames(installed.packages())))

#Load all packages
invisible(lapply(list.of.packages, library, character.only = TRUE)) #use 'invisible' to prevent output being displayed

## Clean workspace & set up environment ----
rm(list=ls()) #Ensure empty workspace if running from beginning
here::i_am("code/06_spatial_graphics.Rmd")
here::here() #Check here location
#OR
#setwd() #Set working directory directly
options(digits = 6) #Set standard decimal print output
options(scipen = 999) #Turn scientific notation on and off (0 = on, 999 = off)

#Ensure necessary directories are created
figsDir <- here::here("figs")
if (!dir.exists(figsDir)){
  dir.create(figsDir)
}
figsDir <- here::here("figs/both_lcms_lcmap")
if (!dir.exists(figsDir)){
  dir.create(figsDir)
}

derivedDatDir <- here::here("data", "derived")
if (!dir.exists(derivedDatDir)){
  dir.create(derivedDatDir)
}

```

# Read in data
Read in the data;
-From Rud: https://docs.google.com/spreadsheets/d/1cGDeCfF6Gzkf9P4SRHDfqPmZfwyX438jURyzrL49IGk/edit?pli=1#gid=0
  (has three tabs of data) - all in HA
-Raw MTBS perimeter data
-NFPORS + GEE extracted data

```{r}
#From Rud: MTBS & Small Fire data
fullDataMTBS <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1cGDeCfF6Gzkf9P4SRHDfqPmZfwyX438jURyzrL49IGk/edit?pli=1#gid=0", sheet = "MTBS Fire Events")
fullDataSmall <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1cGDeCfF6Gzkf9P4SRHDfqPmZfwyX438jURyzrL49IGk/edit?pli=1#gid=0", sheet = "Small Fire Events")

#Raw MTBS perimeter data
mtbsPerims <- sf::st_read(here::here('data', 'raw', 'mtbs_perimeter_data_1984_2021', 'mtbs_perims_DD.shp'))

#Small fire events from Welty, sent to Rud (generated from script 07)
weltySmallPerims <- sf::st_read(here::here('data', 'derived', 'welty_jeffries_western_wildfire_sub_1000ac.gpkg')) |>
  sf::st_transform(sf::st_crs(mtbsPerims)) #reproject to match MTBS

#NFPORS rx data from GEE
nfporsWithGee <- read_csv(here::here(derivedDatDir, "nfpors_with_gee_lcms_lcmap.csv"))

#Raw NFPORS location data
nfpors <- sf::st_read(here::here("data", "raw", "West_NFPORS_2010_2021", "West_NFPORS_2010_2021.shp")) |>
  sf::st_transform(sf::st_crs(mtbsPerims)) #reproject to match MTBS


```
# Prep MTBS data

```{r}

#Join together perimeters and good fire data
mtbsPerimsInterest <- mtbsPerims |>
  dplyr::filter(Event_ID %in% fullDataMTBS$Event_ID) |>
  dplyr::left_join(fullDataMTBS, by = c("Event_ID"))|>
  dplyr::filter(preFireFor > 0) |>
  dplyr::mutate(goodPercFor = round(100 * (frcLowCbiLow / preFireFor)),
                goodPercAll = round(100 * (frcLowCbiLow / areaHA)),
                percFor = round(100 * (preFireFor / areaHA)))

#Get perimeter centroids to use as points
mtbsPointsInterest <- mtbsPerimsInterest |>
  sf::st_make_valid() |> #Make valid first since there is an invalid polygon in the mix
  sf::st_centroid()

```

# Prep Welty data

```{r}

#Join together perimeters and good fire data
weltyInterest <- weltySmallPerims |>
  dplyr::filter(OBJECTID %in% fullDataSmall$OBJECTID) |>
  dplyr::left_join(fullDataSmall, by = c("OBJECTID")) |>
  dplyr::mutate(goodPercAll = round(100 * (frcLowCbiLowHA / GIS_Hectar.y)))

#Get perimeter centroids to use as points
sf_use_s2(FALSE) #Polygon invalidity not fixed by st_make_valid, turn spherical geometry on and off to get centroids
weltyPointsInterest <- weltyInterest |>
  sf::st_make_valid() |> #Make valid first since there is an invalid polygon in the mix
  sf::st_centroid()
sf_use_s2(TRUE)

```

#Merge MTBS & Welty data
Points, perimeters, and forest-filtered perimeters

```{r}

goodFirePoints <- rbind(
  (mtbsPointsInterest |> dplyr::select(frcLowCbiLow)),
  (weltyPointsInterest |>
     dplyr::select(frcLowCbiLowHA) |> #make sure to use HA column
     dplyr::rename("geometry" = "geom",
                   "frcLowCbiLow" = "frcLowCbiLowHA"))
)


goodFirePerims <- rbind(
  (mtbsPerimsInterest |> dplyr::select(frcLowCbiLow, goodPercAll)),
  (weltyInterest |>
     dplyr::select(frcLowCbiLowHA, goodPercAll) |>
     dplyr::rename("geometry" = "geom",
                   "frcLowCbiLow" = "frcLowCbiLowHA"))
)


goodFirePerimsForestOnly <- rbind( #good fire perimeters filtered to only include those with >10% forest pre-burn
  (mtbsPerimsInterest |>
     dplyr::mutate(percPreFor = (preFireFor / areaHA) * 100) |>
     dplyr::filter(percPreFor >= 10) |>
     dplyr::mutate(goodPercBurnedFor = frcLowCbiLow / (CBIhighsev + CBIlowmodsev) * 100) |>
     dplyr::select(frcLowCbiLow, goodPercAll, goodPercBurnedFor, areaHA)),
  (weltyInterest |>
     dplyr::mutate(percPreFor = (preFireFor / GIS_Hectar.y) * 100) |>
     dplyr::filter(percPreFor >= 10) |>
      dplyr::mutate(goodPercBurnedFor = frcLowCbiLowHA / (CBIlowmodsevHA + CBIhighsevHA) * 100) |>
     dplyr::select(frcLowCbiLowHA, goodPercAll, goodPercBurnedFor, GIS_Hectar.y) |>
     dplyr::rename("geometry" = "geom",
                   "frcLowCbiLow" = "frcLowCbiLowHA",
                   "areaHA" = "GIS_Hectar.y"))
)

```

# Prep NFPORS data

```{r}

#Join the NFPORS GEE data back to the original points & filter
nfporsInterest <- nfpors |>
  dplyr::left_join(nfporsWithGee, by = c("PointId" = "objectid__")) |>
  dplyr::filter(FRGDescription == "frcLowMix"& ACTUALCOMPLETIONYEARNEW < 2021) |>
  dplyr::filter(LCMS_PreBurnYear == 1 & LCMAP_PreBurnYear == 4)

```

# Get contextual data

```{r context-dat}
#Pull in state data as context
usa <- tigris::states() |>
  sf::st_transform(sf::st_crs(mtbsPerims))
west <- usa[usa$STUSPS %in% c("WA", "OR", "CA", "ID", "MT", "WY", "NV", "AZ", "CO", "NM", "UT"),]  
```
# Create histogram

```{r}


create.event.summary.plots <- function(eventDats, nm) {
  ggplot() +
  ggplot2::geom_histogram(data = eventDats, aes(x=goodPercBurnedFor), fill = "#dc0ab4ff", bins = 50) +
  ggplot2::labs(x = "Percentage of forest burned in good fire", y = "Count") +
  ggplot2::theme_minimal() +
  scale_y_reverse() +
  coord_flip()
  ggplot2::ggsave(filename = here::here(figsDir, glue::glue('event_summary_hist_{nm}.png')), units = "in", width = 3, height = 3, bg = "white")
  
ggplot() +
  ggplot2::geom_point(data = eventDats, aes(x = areaHA, y = goodPercBurnedFor), color = "#dc0ab4ff", alpha = 0.5, stroke = NA) +
  ggplot2::labs(x = "Area", y = "Percentage of forest burned in good fire") +
  ggplot2::theme_minimal()
  ggplot2::ggsave(filename = here::here(figsDir, glue::glue('event_summary_point_{nm}.png')), units = "in", width = 3, height = 3)
}

largeGoodFirePerimsForestOnly <- goodFirePerimsForestOnly |> dplyr::filter(areaHA > 1000)


create.event.summary.plots(goodFirePerimsForestOnly, "allFires")
create.event.summary.plots(largeGoodFirePerimsForestOnly, "largeFires")



```



# Create spatial map

```{r spatial-map}

# Create a leaflet map to get accurate (geographically meaningful) size of NFPORS points
# Since leaflet does not export easily for static graphics, the information from this has been visually used
# to adjust the GGPLOT2 final outputs
goodFCol = "#dc0ab4ff"
lowGoodFCol = "gray50"
gfBCol = "transparent"
rxCol = "goldenrod1"
rxBCol = "transparent"
transparency = 0.75
strk = 0
rngMin = 0.1
rngMax = 7
pch = 21

library(leaflet)
nfporsInterest2 <- nfporsInterest |>
  dplyr::mutate(area_m2 = TOTALACCOMPLISHMENT_HA * 10000) |> # 1 hectare = 10,000 square meters
  dplyr::mutate(radius = sqrt(area_m2 / pi)) %>%
  cbind(sf::st_coordinates(.)) |>
  dplyr::rename(lat = Y, long = X)

m <- leaflet::leaflet() %>%
  leaflet::addPolygons(data = west, fillOpacity = 0.8, fillColor = "white", color = "#444444") %>%
  leaflet::addPolygons(data = goodFirePerimsForestOnly, fillColor = goodFCol, fillOpacity = 0.8, color = NA) %>%
  leaflet::addCircles(data = nfporsInterest2, lng = ~long, lat = ~lat, radius = ~radius, fillColor = rxCol, color = rxBCol, opacity = transparency, weight = strk, fillOpacity = transparency)

# Display the map
m
```

```{r}
#Function and call to make figure

create.plots <- function(goodFCol, lowGoodFCol, gfBCol, rxCol, rxBCol, transparency, strk, rngMin, rngMax, pch, bgfill, statecol) {
  
  tm <- Sys.time() |>
    gsub(pattern = " ", replacement = "_") |>
    gsub(pattern = ":", replacement = "") |>
    gsub(pattern = ".", replacement = "", fixed = TRUE)
  
  #Create perims + bubbles plot
  ggplot2::ggplot() +
    ggplot2::geom_sf(data = west, aes(), fill = bgfill, color = statecol) +
    #ggplot2::geom_sf(data = goodFirePerimsForestOnly, aes(fill = goodPercAll), color = NA) +
    ggplot2::geom_sf(data = goodFirePerimsForestOnly, aes(fill = goodPercBurnedFor), color = NA) + #use percentage of only burned forest within perimeter
    ggplot2::scale_fill_gradient(low = lowGoodFCol, high = goodFCol, limits = c(0, 100)) +
    ggplot2::geom_sf(data = nfporsInterest, aes(size = TOTALACCOMPLISHMENT_HA), fill = rxCol, color = rxBCol, pch = pch, alpha = transparency, stroke = strk) +
    #ggplot2::labs(title = "Good fire & Rx Burns in the Western US (rx on top, perimeters) \n Perimeters filtered to >10% forest before burn", fill = "Percentage of \ngood forest fire \nwithin perimeter")  +
    ggplot2::labs(
      #title = "Good fire & Rx Burns in the Western US (rx on top, perimeters) \n Perimeters filtered to >10% forest before burn",
      fill = "Percentage of \nforest within \nperimeter \nburned by \ngood fire")  +
    theme_void() +
    theme(plot.background = element_rect(fill = "white"),
          panel.border = element_blank(),
          legend.spacing.y = ggplot2::unit(0.6, 'cm'),
          text = element_text(family = "Arial", size = 12, lineheight = 1.15)) +
    ggplot2::scale_size_continuous(
      labels = scales::label_comma(),
      limits = c(0, max(max(goodFirePoints$frcLowCbiLow), max(nfporsInterest$TOTALACCOMPLISHMENT_HA))),
      name = "Prescribed \nburn \nhectares",
      range = c(rngMin, rngMax))
  ggplot2::ggsave(filename = here::here(figsDir, glue::glue('westwide_forestPerims_rxBubbles_{tm}.png')), units = "in", width = 7)
  
}


create.plots(
  goodFCol = "#dc0ab4ff",
  lowGoodFCol = "goldenrod2",
  gfBCol = "transparent",
  rxCol = "darkblue",
  rxBCol = "transparent",
  transparency = 1,
  strk = 0,
  rngMin = 0.1,
  rngMax = 2.65, #tested until visual comparison of bubbles to that from leaflet is as similar as possible
  pch = 21,
  bgfill = "gray95",
  statecol = "gray50"
)

# Function to get a hex color from R color
# PARAMETERS
# color : a standard R color (e.g. 'red' or 'steelblue')
col2hex <- function(color) {
  rgb_values <- col2rgb(color)
  hex_color <- rgb(rgb_values[1], rgb_values[2], rgb_values[3], maxColorValue=255)
  return(hex_color)
}
col2hex("goldenrod2")
col2hex("darkblue")
col2hex("gray95")
col2hex("gray50")



```

# OLD FIGURES

```{r}
# ####################################################################################
# #FIRST SET
# 
# #goodFCol = "goldenrod2"
# goodFCol = "#dc0ab4ff"
# #rxCol = "seagreen"
# rxCol = "goldenrod2"
# transparency = 0.6
# strk = 0.3
# rngMax = 5
# 
# #Create plot with both on one
# #rx on top
# ggplot2::ggplot() +
#   ggplot2::geom_sf(data = west, aes(), fill = "white") +
#   ggplot2::geom_sf(data = goodFirePoints, aes(size = frcLowCbiLow), fill = goodFCol, color = goodFCol, pch = 21, alpha = transparency, stroke = strk) +
#   ggplot2::geom_sf(data = nfporsInterest, aes(size = TOTALACCOMPLISHMENT_HA), fill = rxCol, color = rxCol, pch = 21, alpha = transparency, stroke = strk) +
#   ggplot2::labs(title = "Area burned in 'good fire' and prescribed fire events \n in the Western US (rx on top)") +
#   ggplot2::scale_size_continuous(limits = c(0, max(max(goodFirePoints$frcLowCbiLow), max(nfporsInterest$TOTALACCOMPLISHMENT_HA))),
#                                  name = "Hectares",
#                                  range = c(0, rngMax)) +
#   theme_bw()
# ggplot2::ggsave(filename = here::here(figsDir, 'spatial_plot_1.png'), units = "in", width = 7)
# 
# #rx on top v2 - make everything even smaller by forcing plot to larger width/height
# rngMax = 3
# ggplot2::ggplot() +
#   ggplot2::geom_sf(data = west, aes(), fill = "white") +
#   ggplot2::geom_sf(data = goodFirePoints, aes(size = frcLowCbiLow), fill = goodFCol, color = goodFCol, pch = 21, alpha = transparency, stroke = strk) +
#   ggplot2::geom_sf(data = nfporsInterest, aes(size = TOTALACCOMPLISHMENT_HA), fill = rxCol, color = rxCol, pch = 21, alpha = transparency, stroke = strk) +
#   ggplot2::labs(title = "Area burned in 'good fire' and prescribed fire events \n in the Western US (rx on top)") +
#   ggplot2::scale_size_continuous(limits = c(0, max(max(goodFirePoints$frcLowCbiLow), max(nfporsInterest$TOTALACCOMPLISHMENT_HA))),
#                                  name = "Hectares",
#                                  range = c(0, rngMax)) +
#   theme_bw()
# ggplot2::ggsave(filename = here::here(figsDir, 'spatial_plot_1v2.png'), units = "in", width = 7)
# 
# #gf on top
# ggplot2::ggplot() +
#   ggplot2::geom_sf(data = west, aes(), fill = "white") +
#     ggplot2::geom_sf(data = nfporsInterest, aes(size = TOTALACCOMPLISHMENT_HA), fill = rxCol, color = rxCol, pch = 21, alpha = transparency, stroke = strk) +
#   ggplot2::geom_sf(data = goodFirePoints, aes(size = frcLowCbiLow), fill = goodFCol, color = goodFCol, pch = 21, alpha = transparency, stroke = strk) +
#   ggplot2::labs(title = "Good fire & Rx Burns in the Western US (fire on top)") +
#   ggplot2::scale_size_continuous(
#     limits = c(0, max(max(goodFirePoints$frcLowCbiLow), max(nfporsInterest$TOTALACCOMPLISHMENT_HA))),
#     name = "Hectares",
#     range = c(0, rngMax)) +
#   theme_bw()
# ggplot2::ggsave(filename = here::here(figsDir, 'spatial_plot_2.png'), units = "in", width = 7)
# 
# 
# #Create plot with two separate
# goodFirePlot <- ggplot2::ggplot() +
#   ggplot2::geom_sf(data = west, aes(), fill = "white") +
#   ggplot2::geom_sf(data = goodFirePoints, aes(size = frcLowCbiLow), fill = goodFCol, color = goodFCol, pch = 21, alpha = transparency, stroke = strk) +
#   ggplot2::scale_size_continuous(limits = c(0, max(max(goodFirePoints$frcLowCbiLow), max(nfporsInterest$TOTALACCOMPLISHMENT_HA))),
#                                  name = "Good Fire\nHa",
#                                  range = c(0, rngMax)) +
#   theme_bw()
# 
# rxFirePlot <- ggplot2::ggplot() +
#   ggplot2::geom_sf(data = west, aes(), fill = "white") +
#   ggplot2::geom_sf(data = nfporsInterest, aes(size = TOTALACCOMPLISHMENT_HA), fill = rxCol, color = rxCol, pch = 21, alpha = transparency, stroke = strk) +
#   ggplot2::scale_size_continuous(limits = c(0, max(max(goodFirePoints$frcLowCbiLow), max(nfporsInterest$TOTALACCOMPLISHMENT_HA))),
#                                  name = "Rx Ha",
#                                  range = c(0, rngMax)) +
#   theme_bw()
# 
# combinedPlot <- gridExtra::arrangeGrob(goodFirePlot, rxFirePlot, ncol = 2, top = "Good Fire & Rx in the Western US")
# ggplot2::ggsave(combinedPlot, filename = here::here(figsDir, 'spatial_plot_3.png'), units = "in", width = 10)
# 
# 
# ########################################################################################
# # PERIMETER SET
# 
# #darkorchid3
# goodFCol = "#dc0ab4ff"
# 
# goodFirePerims <- rbind(
#   (mtbsPerimsInterest |> dplyr::select(frcLowCbiLow, goodPercAll)),
#   (weltyInterest |>
#      dplyr::select(frcLowCbiLowHA, goodPercAll) |>
#      dplyr::rename("geometry" = "geom",
#                    "frcLowCbiLow" = "frcLowCbiLowHA"))
# )
# 
# 
# goodFirePerimsForestOnly <- rbind( #good fire perimeters filtered to only include those with >10% forest pre-burn
#   (mtbsPerimsInterest |>
#      dplyr::mutate(percPreFor = (preFireFor / areaHA) * 100) |>
#      dplyr::filter(percPreFor >= 10) |>
#      dplyr::select(frcLowCbiLow, goodPercAll)),
#   (weltyInterest |>
#      dplyr::mutate(percPreFor = (preFireFor / GIS_Hectar.y) * 100) |>
#      dplyr::filter(percPreFor >= 10) |>
#      dplyr::select(frcLowCbiLowHA, goodPercAll) |>
#      dplyr::rename("geometry" = "geom",
#                    "frcLowCbiLow" = "frcLowCbiLowHA"))
# )
# 
# #Create graphic 1
# ggplot2::ggplot() +
#   ggplot2::geom_sf(data = west, aes(), fill = "white") +
#   ggplot2::geom_sf(data = goodFirePerims, aes(fill = goodPercAll), color = NA) +
#   ggplot2::scale_fill_gradient(low = "gray", high = goodFCol, limits = c(0, 100)) +
#   ggplot2::geom_sf(data = nfporsInterest, size = 0.0001, fill = rxCol, color = rxCol, pch = 21, alpha = transparency, stroke = strk) +
#   ggplot2::labs(title = "Good fire & Rx Burns in the Western US (rx on top, perimeters)", fill = "Percentage of \ngood forest fire \nwithin perimeter")  +
#   theme_bw()
# ggplot2::ggsave(filename = here::here(figsDir, 'spatial_plot_perims.png'), units = "in", width = 7)
# 
# 
# #make everything even smaller by forcing plot to larger width/height
# ggplot2::ggplot() +
#   ggplot2::geom_sf(data = west, aes(), fill = "white") +
#   ggplot2::geom_sf(data = goodFirePerims, aes(fill = goodPercAll), color = NA) +
#   ggplot2::scale_fill_gradient(low = "gray", high = goodFCol, limits = c(0, 100)) +
#   ggplot2::geom_sf(data = nfporsInterest, size = 0.0001, fill = rxCol, color = rxCol, pch = 21, alpha = transparency, stroke = strk) +
#   ggplot2::labs(title = "Good fire & Rx Burns in the Western US (rx on top, perimeters)", fill = "Percentage of \ngood forest fire \nwithin perimeter")  +
#   theme_bw()
# ggplot2::ggsave(filename = here::here(figsDir, 'spatial_plot_perims2.png'), units = "in", width = 10, height = 6.5)
# 
# 
# #create graphic 2
# ggplot2::ggplot() +
#   ggplot2::geom_sf(data = west, aes(), fill = "white") +
#   ggplot2::geom_sf(data = goodFirePerims, aes(fill = goodPercAll), color = NA) +
#   ggplot2::scale_fill_gradient(low = "gray", high = goodFCol, limits = c(0, 100)) +
#   ggplot2::labs(title = "Good fire in the Western US (perimeters)", fill = "Percentage of \ngood forest fire \nwithin perimeter")  +
#   theme_bw()
# ggplot2::ggsave(filename = here::here(figsDir, 'spatial_plot_only_perims.png'), units = "in", width = 7)
# 
# 
# 
# #Create plot with two separate
# goodFirePlot2 <- ggplot2::ggplot() +
#   ggplot2::geom_sf(data = west, aes(), fill = "white") +
#   ggplot2::geom_sf(data = goodFirePerims, aes(fill = goodPercAll), color = NA) +
#   ggplot2::scale_fill_gradient(low = "gray", high = goodFCol, limits = c(0, 100)) +
#   theme_bw()
# 
# rxFirePlot2 <- ggplot2::ggplot() +
#   ggplot2::geom_sf(data = west, aes(), fill = "white") +
#   ggplot2::geom_sf(data = nfporsInterest, size = 0.0001, fill = rxCol, color = rxCol, pch = 21, alpha = transparency, stroke = strk) +
#   theme_bw()
# 
# combinedPlot <- gridExtra::arrangeGrob(goodFirePlot2, rxFirePlot2, ncol = 2, top = "Good Fire & Rx in the Western US")
# ggplot2::ggsave(combinedPlot, filename = here::here(figsDir, 'spatial_plot_2_perims.png'), units = "in", width = 10)
# 
# 
# 
# 
# #FOREST ONLY
# 
# 
# #Create graphic 1
# ggplot2::ggplot() +
#   ggplot2::geom_sf(data = west, aes(), fill = "white") +
#   ggplot2::geom_sf(data = goodFirePerimsForestOnly, aes(fill = goodPercAll), color = NA) +
#   ggplot2::scale_fill_gradient(low = "gray", high = goodFCol, limits = c(0, 100)) +
#   ggplot2::geom_sf(data = nfporsInterest, size = 0.0001, fill = rxCol, color = rxCol, pch = 21, alpha = transparency, stroke = strk) +
#   ggplot2::labs(title = "Good fire & Rx Burns in the Western US (rx on top, perimeters) : >10% forest within perimeters", fill = "Percentage of \ngood forest fire \nwithin perimeter")  +
#   theme_bw()
# ggplot2::ggsave(filename = here::here(figsDir, 'spatial_plot_perims_forestOnly.png'), units = "in", width = 7)
# 
# #make everything even smaller by forcing plot to larger width/height
# ggplot2::ggplot() +
#   ggplot2::geom_sf(data = west, aes(), fill = "white") +
#   ggplot2::geom_sf(data = goodFirePerimsForestOnly, aes(fill = goodPercAll), color = NA) +
#   ggplot2::scale_fill_gradient(low = "gray", high = goodFCol, limits = c(0, 100)) +
#   ggplot2::geom_sf(data = nfporsInterest, size = 0.0001, fill = rxCol, color = rxCol, pch = 21, alpha = transparency, stroke = strk) +
#   ggplot2::labs(title = "Good fire & Rx Burns in the Western US (rx on top, perimeters) : >10% forest within perimeters", fill = "Percentage of \ngood forest fire \nwithin perimeter")  +
#   theme_bw()
# ggplot2::ggsave(filename = here::here(figsDir, 'spatial_plot_perims2_forestOnly.png'), units = "in", width = 10, height = 6.5)
# 
# 
# #create graphic 2
# ggplot2::ggplot() +
#   ggplot2::geom_sf(data = west, aes(), fill = "white") +
#   ggplot2::geom_sf(data = goodFirePerimsForestOnly, aes(fill = goodPercAll), color = NA) +
#   ggplot2::scale_fill_gradient(low = "gray", high = goodFCol, limits = c(0, 100)) +
#   ggplot2::labs(title = "Good fire in the Western US (perimeters)", fill = "Percentage of \ngood forest fire \nwithin perimeter : >10% forest within perimeters")  +
#   theme_bw()
# ggplot2::ggsave(filename = here::here(figsDir, 'spatial_plot_only_perims_forestOnly.png'), units = "in", width = 7)
# 
# 
# 
# #Create plot with two separate
# goodFirePlot2 <- ggplot2::ggplot() +
#   ggplot2::geom_sf(data = west, aes(), fill = "white") +
#   ggplot2::geom_sf(data = goodFirePerimsForestOnly, aes(fill = goodPercAll), color = NA) +
#   ggplot2::scale_fill_gradient(low = "gray", high = goodFCol, limits = c(0, 100)) +
#   theme_bw()
# 
# rxFirePlot2 <- ggplot2::ggplot() +
#   ggplot2::geom_sf(data = west, aes(), fill = "white") +
#   ggplot2::geom_sf(data = nfporsInterest, size = 0.0001, fill = rxCol, color = rxCol, pch = 21, alpha = transparency, stroke = strk) +
#   theme_bw()
# 
# combinedPlot <- gridExtra::arrangeGrob(goodFirePlot2, rxFirePlot2, ncol = 2, top = "Good Fire & Rx in the Western US : >10% forest within perimeters")
# ggplot2::ggsave(combinedPlot, filename = here::here(figsDir, 'spatial_plot_2_perims_forestOnly.png'), units = "in", width = 10)
# 

############################################
# OLD FUNCTION
# 
# 
# create.plots <- function(goodFCol, lowGoodFCol, gfBCol, rxCol, rxBCol, transparency, strk, rngMin, rngMax, pch, bgfill, statecol) {
#   
#   tm <- Sys.time() |>
#     gsub(pattern = " ", replacement = "_") |>
#     gsub(pattern = ":", replacement = "") |>
#     gsub(pattern = ".", replacement = "", fixed = TRUE)
#   
#   # #Create bubble plot
#   # ggplot2::ggplot() +
#   #   ggplot2::geom_sf(data = west, aes(), fill = "white") +
#   #   ggplot2::geom_sf(data = goodFirePoints, aes(size = frcLowCbiLow), fill = goodFCol, color = gfBCol, pch = pch, alpha = transparency, stroke = strk) +
#   #   ggplot2::geom_sf(data = nfporsInterest, aes(size = TOTALACCOMPLISHMENT_HA), fill = rxCol, color = rxBCol, pch = pch, alpha = transparency, stroke = strk) +
#   #   ggplot2::labs(title = "Area burned in 'good fire' and prescribed fire events \n in the Western US (rx on top)") +
#   #   ggplot2::scale_size_continuous(limits = c(0, max(max(goodFirePoints$frcLowCbiLow), max(nfporsInterest$TOTALACCOMPLISHMENT_HA))),
#   #                                  name = "Hectares",
#   #                                  range = c(rngMin, rngMax)) +
#   #   theme_bw()
#   # ggplot2::ggsave(filename = here::here(figsDir, glue::glue('westwide_bubbles_{tm}.png')), units = "in", width = 7)
#   
#   #Create perims + bubbles plot
#   ggplot2::ggplot() +
#     ggplot2::geom_sf(data = west, aes(), fill = bgfill, color = statecol) +
#     #ggplot2::geom_sf(data = goodFirePerimsForestOnly, aes(fill = goodPercAll), color = NA) +
#     ggplot2::geom_sf(data = goodFirePerimsForestOnly, aes(fill = goodPercBurnedFor), color = NA) + #use percentage of only burned forest within perimeter
#     ggplot2::scale_fill_gradient(low = lowGoodFCol, high = goodFCol, limits = c(0, 100)) +
#     ggplot2::geom_sf(data = nfporsInterest, aes(size = TOTALACCOMPLISHMENT_HA), fill = rxCol, color = rxBCol, pch = pch, alpha = transparency, stroke = strk) +
#     #ggplot2::labs(title = "Good fire & Rx Burns in the Western US (rx on top, perimeters) \n Perimeters filtered to >10% forest before burn", fill = "Percentage of \ngood forest fire \nwithin perimeter")  +
#     ggplot2::labs(
#       #title = "Good fire & Rx Burns in the Western US (rx on top, perimeters) \n Perimeters filtered to >10% forest before burn",
#       fill = "Percentage of \nforest within \nperimeter \nburned by \ngood fire")  +
#     theme_void() +
#     theme(plot.background = element_rect(fill = "white"),
#           panel.border = element_blank(),
#           legend.spacing.y = ggplot2::unit(1, 'cm'),
#           text = element_text(family = "Arial", size = 10, lineheight = 1.2)) +
#     ggplot2::scale_size_continuous(
#       labels = scales::label_comma(),
#       limits = c(0, max(max(goodFirePoints$frcLowCbiLow), max(nfporsInterest$TOTALACCOMPLISHMENT_HA))),
#       name = "Prescribed \nburn \nhectares",
#       range = c(rngMin, rngMax))
#   ggplot2::ggsave(filename = here::here(figsDir, glue::glue('westwide_forestPerims_rxBubbles_{tm}.png')), units = "in", width = 7)
#   
# }

```

