Create spatial graphics for good fire
Tyler L. McIntosh
CU Boulder CIRES Earth Lab
Last updated: 12/12/23

This script uses the following naming conventions wherever possible:
 lowerCamelCase for variables
 period.separated for functions
 underscore_separated for files

# Setup

```{r, message = FALSE, warning = FALSE, include = FALSE}
## Libraries ----

#Check the required libraries and download if needed
list.of.packages <- c("tidyverse", #Includes ggplot2, dplyr, tidyr, readr, purrr, tibble, stringr, forcats
                      "terra",
                      "sf",
                      "mapview",
                      "here",
                      "future", "future.apply", "furrr", "doFuture", "progressr", #Futureverse!
                      "tictoc", 
                      "mblm", #Median-based linear models (i.e. thiel-sen)
                      "plyr",
                      "kableExtra", #nice tables
                      "googlesheets4", #read in google sheet data
                      "scales", #add commas to ggplot axis
                      "tigris", #US data
                      "tmap", #mapping
                      "gridExtra", #multi-frame plots
                      "scales") #add commas to axis

#Install all packages that aren't installed yet
install.packages(setdiff(list.of.packages, rownames(installed.packages())))

#Load all packages
invisible(lapply(list.of.packages, library, character.only = TRUE)) #use 'invisible' to prevent output being displayed

## Clean workspace & set up environment ----
rm(list=ls()) #Ensure empty workspace if running from beginning
here::i_am("code/06_spatial_graphics.Rmd")
here::here() #Check here location
#OR
#setwd() #Set working directory directly
options(digits = 6) #Set standard decimal print output
options(scipen = 999) #Turn scientific notation on and off (0 = on, 999 = off)

#Ensure necessary directories are created
figsDir <- here::here("figs")
if (!dir.exists(figsDir)){
  dir.create(figsDir)
}

derivedDatDir <- here::here("data", "derived")
if (!dir.exists(derivedDatDir)){
  dir.create(derivedDatDir)
}

```

# Read in data
Read in the data;
-From Rud: https://docs.google.com/spreadsheets/d/1cGDeCfF6Gzkf9P4SRHDfqPmZfwyX438jURyzrL49IGk/edit?pli=1#gid=0
  (has three tabs of data) - all in HA
-Raw MTBS perimeter data
-NFPORS + GEE extracted data

```{r}
#From script 03 - all 3 sheets of data
fullDataMTBS <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1cGDeCfF6Gzkf9P4SRHDfqPmZfwyX438jURyzrL49IGk/edit?pli=1#gid=0", sheet = "Full Data")

#Raw MTBS perimeter data
mtbsPerims <- sf::st_read(here::here('data', 'raw', 'mtbs_perimeter_data_1984_2021', 'mtbs_perims_DD.shp'))

#NFPORS rx data from GEE
nfporsWithGee <- read_csv(here::here(derivedDatDir, "nfpors_with_gee.csv"))

#Raw NFPORS location data
nfpors <- sf::st_read(here::here("data", "raw", "West_NFPORS_2010_2021", "West_NFPORS_2010_2021.shp")) |>
  sf::st_transform(sf::st_crs(mtbsPerims)) #reproject to match MTBS


```
# Prep MTBS data

```{r}

#Join together perimeters and good fire data
mtbsPerimsInterest <- mtbsPerims |>
  dplyr::filter(Event_ID %in% fullDataMTBS$Event_ID) |>
  dplyr::left_join(fullDataMTBS, by = c("Event_ID"))|>
  dplyr::filter(preFireFor > 0) |>
  dplyr::mutate(goodPercFor = round(100 * (frcLowCbiLow / preFireFor)),
                goodPercAll = round(100 * (frcLowCbiLow / areaHA)),
                percFor = round(100 * (preFireFor / areaHA)))

#Get perimeter centroids to use as points
mtbsPointsInterest <- mtbsPerimsInterest |>
  sf::st_make_valid() |> #Make valid first since there is an invalid polygon in the mix
  sf::st_centroid()

```

# Prep NFPORS data

```{r}

#Join the NFPORS GEE data back to the original points
nfporsInterest <- nfpors |>
  dplyr::left_join(nfporsWithGee, by = c("PointId" = "objectid__")) |>
  dplyr::filter(FRGDescription == "frcLowMix")

```

# Get contextual data

```{r}
#Pull in state data as context
usa <- tigris::states() |>
  sf::st_transform(sf::st_crs(mtbsPerims))
west <- usa[usa$STUSPS %in% c("WA", "OR", "CA", "ID", "MT", "WY", "NV", "AZ", "CO", "NM", "UT"),]  
```


# Create spatial map

```{r}
goodFCol = "goldenrod2"
rxCol = "seagreen"
transparency = 0.6
strk = 0.3
rngMax = 5

#Create plot with both on one
ggplot2::ggplot() +
  ggplot2::geom_sf(data = west, aes(), fill = "white") +
  ggplot2::geom_sf(data = mtbsPointsInterest, aes(size = frcLowCbiLow), fill = goodFCol, color = goodFCol, pch = 21, alpha = transparency, stroke = strk) +
  ggplot2::geom_sf(data = nfporsInterest, aes(size = TOTALACCOMPLISHMENT_HA), fill = rxCol, color = rxCol, pch = 21, alpha = transparency, stroke = strk) +
  ggplot2::labs(title = "Good fire & Rx Burns in the Western US (rx on top)") +
  ggplot2::scale_size_continuous(limits = c(0, max(max(mtbsPointsInterest$frcLowCbiLow), max(nfporsInterest$TOTALACCOMPLISHMENT_HA))),
                                 name = "Hectares",
                                 range = c(0, rngMax)) +
  theme_bw()
ggplot2::ggsave(filename = here::here(figsDir, 'spatial_plot_1.png'), units = "in", width = 7)

ggplot2::ggplot() +
  ggplot2::geom_sf(data = west, aes(), fill = "white") +
    ggplot2::geom_sf(data = nfporsInterest, aes(size = TOTALACCOMPLISHMENT_HA), fill = rxCol, color = rxCol, pch = 21, alpha = transparency, stroke = strk) +
  ggplot2::geom_sf(data = mtbsPointsInterest, aes(size = frcLowCbiLow), fill = goodFCol, color = goodFCol, pch = 21, alpha = transparency, stroke = strk) +
  ggplot2::labs(title = "Good fire & Rx Burns in the Western US (fire on top)") +
  ggplot2::scale_size_continuous(limits = c(0, max(max(mtbsPointsInterest$frcLowCbiLow), max(nfporsInterest$TOTALACCOMPLISHMENT_HA))),
                                 name = "Hectares",
                                 range = c(0, rngMax)) +
  theme_bw()
ggplot2::ggsave(filename = here::here(figsDir, 'spatial_plot_1v2.png'), units = "in", width = 7)


#Create plot with two separate
goodFirePlot <- ggplot2::ggplot() +
  ggplot2::geom_sf(data = west, aes(), fill = "white") +
  ggplot2::geom_sf(data = mtbsPointsInterest, aes(size = frcLowCbiLow), fill = goodFCol, color = goodFCol, pch = 21, alpha = transparency, stroke = strk) +
  ggplot2::scale_size_continuous(limits = c(0, max(max(mtbsPointsInterest$frcLowCbiLow), max(nfporsInterest$TOTALACCOMPLISHMENT_HA))),
                                 name = "Good Fire\nHa",
                                 range = c(0, rngMax)) +
  theme_bw()

rxFirePlot <- ggplot2::ggplot() +
  ggplot2::geom_sf(data = west, aes(), fill = "white") +
  ggplot2::geom_sf(data = nfporsInterest, aes(size = TOTALACCOMPLISHMENT_HA), fill = rxCol, color = rxCol, pch = 21, alpha = transparency, stroke = strk) +
  ggplot2::scale_size_continuous(limits = c(0, max(max(mtbsPointsInterest$frcLowCbiLow), max(nfporsInterest$TOTALACCOMPLISHMENT_HA))),
                                 name = "Rx Ha",
                                 range = c(0, rngMax)) +
  theme_bw()

combinedPlot <- gridExtra::arrangeGrob(goodFirePlot, rxFirePlot, ncol = 2, top = "Good Fire & Rx in the Western US")
ggplot2::ggsave(combinedPlot, filename = here::here(figsDir, 'spatial_plot_2.png'), units = "in", width = 10)


```





