Examine potential datasets for small fires

Tyler L. McIntosh
CU Boulder CIRES Earth Lab
Last updated: 12/14/23

This script uses the following naming conventions wherever possible:
 lowerCamelCase for variables
 period.separated for functions
 underscore_separated for files

# Setup

```{r, message = FALSE, warning = FALSE, include = FALSE}
## Libraries ----

#Check the required libraries and download if needed
list.of.packages <- c("tidyverse", #Includes ggplot2, dplyr, tidyr, readr, purrr, tibble, stringr, forcats
                      "terra",
                      "sf",
                      "mapview",
                      "here",
                      "future", "future.apply", "furrr", "doFuture", "progressr", #Futureverse!
                      "tictoc", 
                      "mblm", #Median-based linear models (i.e. thiel-sen)
                      "plyr",
                      "kableExtra", #nice tables
                      "googlesheets4", #read in google sheet data
                      "tigris", #US data
                      "tmap", #mapping
                      "gridExtra", #multi-frame plots
                      "scales") #add commas to axis

#Install all packages that aren't installed yet
install.packages(setdiff(list.of.packages, rownames(installed.packages())))

#Load all packages
invisible(lapply(list.of.packages, library, character.only = TRUE)) #use 'invisible' to prevent output being displayed

## Clean workspace & set up environment ----
rm(list=ls()) #Ensure empty workspace if running from beginning
here::i_am("code/06_spatial_graphics.Rmd")
here::here() #Check here location
#OR
#setwd() #Set working directory directly
options(digits = 6) #Set standard decimal print output
options(scipen = 999) #Turn scientific notation on and off (0 = on, 999 = off)

#Ensure necessary directories are created
figsDir <- here::here("figs")
if (!dir.exists(figsDir)){
  dir.create(figsDir)
}

derivedDatDir <- here::here("data", "derived")
if (!dir.exists(derivedDatDir)){
  dir.create(derivedDatDir)
}

```

# Read in data

```{r}
welty <- sf::st_read(here::here('data', 'raw', 'welty_combined_wildland_fire_dataset', 'welty_combined_wildland_fire_perimeters.shp'))
fired <- sf::st_read(here::here('data', 'raw', 'fired_uscan_to2021121_events.gpkg')) |>
  sf::st_transform(sf::st_crs(welty))
fullDataMTBS <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1cGDeCfF6Gzkf9P4SRHDfqPmZfwyX438jURyzrL49IGk/edit?pli=1#gid=0", sheet = "Full Data")
nfporsWithGee <- read_csv(here::here(derivedDatDir, "nfpors_with_gee.csv"))

```
# Filter & split

```{r}

#Pull in state data as context
usa <- tigris::states() |>
  sf::st_transform(sf::st_crs(welty))
west <- usa[usa$STUSPS %in% c("WA", "OR", "CA", "ID", "MT", "WY", "NV", "AZ", "CO", "NM", "UT"),] 

#filter to time period and geographic area of interest, add size category
weltyInterest <- welty |>
  dplyr::filter(Fire_Year >= 2010 & Fire_Year <= 2021) |>
  sf::st_filter(west) |>
  dplyr::mutate(SizeCategory = dplyr::case_when(GIS_Acres >= 1000 ~ "Large",
                                                GIS_Acres < 1000 ~ "Small"))
firedInterest <- fired |>
  dplyr::filter(ig_year >= 2010 & ig_year <= 2021) |>
  sf::st_filter(west) |>
  dplyr::mutate(tot_ar_ha = tot_ar_km2 * 100) |> #convert to ha
  dplyr::mutate(SizeCategory = dplyr::case_when(tot_ar_ha >= 404.686 ~ "Large",  #1000 acres in ha
                                                tot_ar_ha < 404.686 ~ "Small"))

nfporsInterest <- nfporsWithGee |>
  #dplyr::filter(LandCoverDescription == "Trees") |>
  dplyr::mutate(SizeCategory = dplyr::case_when(TOTALACCOMPLISHMENT_HA >= 404.686 ~ "Large",  #1000 acres in ha
                                                TOTALACCOMPLISHMENT_HA < 404.686 ~ "Small"))

#Split welty into RX & wildfire
weltyInterestRx <- weltyInterest |>
  dplyr::filter(Assigned_F == "Prescribed Fire" | Assigned_F == "Unknown - Likely Prescribed Fire")
weltyInterestWF <- weltyInterest|>
  dplyr::filter(Assigned_F == "Wildfire" | Assigned_F == "Likely Wildfire")


```

# Deprecated 

```{r}

weltyInterestWF1000 <- weltyInterestWF |>
  dplyr::filter(SizeCategory == "Large")
firedInterest1000 <- firedInterest |>
  dplyr::filter(SizeCategory == "Large")

weltyInterestWFSub1000 <- weltyInterestWF |>
  dplyr::filter(SizeCategory == "Small")
firedInterestSub1000 <- firedInterest |>
  dplyr::filter(SizeCategory == "Small")


glue::glue("Fires greater than 1,000 acres in the western US in our time period of interest: \n {scales::comma(nrow(weltyInterestWF1000))} fires in the Welty dataset ({scales::comma(round(sum(weltyInterestWF1000$GIS_Hectar)))} total ha) \n {scales::comma(nrow(firedInterest1000))} in the FIRED dataset ({scales::comma(round(sum(firedInterest1000$tot_ar_ha)))} total ha) \n {scales::comma(nrow(fullDataMTBS))} fires in the MTBS dataset ({scales::comma(round(sum(fullDataMTBS$areaHA)))} total ha)")

glue::glue("Fires less than 1,000 acres in the western US in our time period of interest: \n {scales::comma(nrow(weltyInterestWFSub1000))} fires in the Welty dataset ({scales::comma(round(sum(weltyInterestWFSub1000$GIS_Hectar)))} total ha) \n {scales::comma(nrow(firedInterestSub1000))} in the FIRED dataset ({scales::comma(round(sum(firedInterestSub1000$tot_ar_ha)))} total ha)")

glue::glue("Rx burns less than 1,000 acres in the western US in our time period of interest: \n {scales::comma(nrow(dplyr::filter(nfporsInterest, SizeCategory == 'Small')))} fires in the NFPORS dataset \n ({scales::comma(round(sum(dplyr::filter(nfporsInterest, SizeCategory == 'Small')$)))} total ha)")

```

#Create combined simple dataframes

```{r}

#Wildfire
simpleWFDat <- rbind(
  (weltyInterestWF |>
    sf::st_drop_geometry() |>
    dplyr::select(GIS_Hectar, Fire_Year, SizeCategory) |>
    dplyr::mutate(Dataset = "Welty") |>
     `names<-`(c("AreaHA", "Year", "SizeCategory", "Dataset"))),
  (firedInterest |>
    sf::st_drop_geometry() |>
    dplyr::select(tot_ar_ha, ig_year, SizeCategory) |>
    dplyr::mutate(Dataset = "FIRED") |>
     `names<-`(c("AreaHA", "Year", "SizeCategory", "Dataset")))
)
#Prescribed burning
simpleRXDat <- rbind(
  (weltyInterestRx |>
    sf::st_drop_geometry() |>
    dplyr::select(GIS_Hectar, Fire_Year, SizeCategory) |>
    dplyr::mutate(Dataset = "Welty") |>
     `names<-`(c("AreaHA", "Year", "SizeCategory", "Dataset"))),
  (nfporsInterest |>
    sf::st_drop_geometry() |>
    dplyr::select(TOTALACCOMPLISHMENT_HA, ACTUALCOMPLETIONYEARNEW, SizeCategory) |>
    dplyr::mutate(Dataset = "NFPORS") |>
     `names<-`(c("AreaHA", "Year", "SizeCategory", "Dataset")))
)

#Prescribed burning broadcast only
simpleRXDat2 <- rbind(
  (weltyInterestRx |>
    sf::st_drop_geometry() |>
    dplyr::select(GIS_Hectar, Fire_Year, SizeCategory) |>
    dplyr::mutate(Dataset = "Welty") |>
     `names<-`(c("AreaHA", "Year", "SizeCategory", "Dataset"))),
  (nfporsInterest |>
    sf::st_drop_geometry() |>
    dplyr::filter(TYPENAME == "Broadcast Burn") |>
    dplyr::select(TOTALACCOMPLISHMENT_HA, ACTUALCOMPLETIONYEARNEW, SizeCategory) |>
    dplyr::mutate(Dataset = "NFPORS") |>
     `names<-`(c("AreaHA", "Year", "SizeCategory", "Dataset")))
)


smallRx <- sum(filter(simpleRXDat, Dataset == "NFPORS" & SizeCategory == "Small")$AreaHA)
allRx <- sum(filter(simpleRXDat, Dataset == "NFPORS")$AreaHA)
percRxSmall = (smallRx / allRx) * 100
glue::glue("{percRxSmall}% of all rx burns are small in NFPORS dataset")

smallRxB <- sum(filter(simpleRXDat2, Dataset == "NFPORS" & SizeCategory == "Small")$AreaHA)
allRxB <- sum(filter(simpleRXDat2, Dataset == "NFPORS")$AreaHA)
percRxSmallB = (smallRxB / allRxB) * 100
glue::glue("{percRxSmallB}% of all broadcast rx burns are small in NFPORS dataset")





```


# Create graphical summaries of data

```{r}

ggplot2::ggplot(simpleRXDat) +
  ggplot2::geom_col(aes(x = Dataset, y = AreaHA)) +
  ggplot2::facet_wrap(~SizeCategory) +
  ggplot2::scale_y_continuous(labels=scales::comma)
  
ggplot2::ggplot(simpleRXDat2) +
  ggplot2::geom_col(aes(x = Dataset, y = AreaHA)) +
  ggplot2::facet_wrap(~SizeCategory) +
  ggplot2::scale_y_continuous(labels=scales::comma)
  

ggplot2::ggplot(simpleWFDat) +
  ggplot2::geom_col(aes(x = Dataset, y = AreaHA)) +
  ggplot2::facet_wrap(~SizeCategory) +
  ggplot2::scale_y_continuous(labels=scales::comma)



```




# Create wildfire histograms (take 1)

```{r}

#Wildfire small area histogram
ggplot2::ggplot(simpleWFDat |> dplyr::filter(SizeCategory == "Small")) +
  ggplot2::geom_histogram(aes(x = AreaHA, fill = Dataset), alpha = 0.4, bins = 50, position = "identity") +
  ggplot2::labs(title = "FIRED & Welty datasets fire area histogram: \nFires < 1000 acres (405 ha) in W.US from 2010 onwards (50 bins)",
                caption = "A MODIS pixel (500m^2) is 25 ha. This fact drives the FIRED dataset spikes every 25 ha")
ggplot2::ggsave(filename = here::here(figsDir, "fired_welty_areahist.png"), bg = "white")


#Wildfire small year chart
ggplot2::ggplot(simpleWFDat |> dplyr::filter(SizeCategory == "Small")) +
  ggplot2::geom_bar(aes(x = Year, fill = Dataset), alpha = 0.4, position = "dodge") +
  ggplot2::labs(title = "FIRED & Welty datasets fire year plot: \nFires < 1000 acres (405 ha) in W.US from 2010 onwards")
ggplot2::ggsave(filename = here::here(figsDir, "fired_welty_yearplot.png"), bg = "white")

#Wildfire small area histogram year facet
ggplot2::ggplot(simpleWFDat |> dplyr::filter(SizeCategory == "Small")) +
  ggplot2::geom_histogram(aes(x = AreaHA, fill = Dataset), alpha = 0.4, bins = 50, position = "identity") +
  ggplot2::labs(title = "FIRED & Welty datasets fire area histogram: \nFires < 1000 acres (405 ha) in W.US from 2010 onwards (50 bins)",
                caption = "A MODIS pixel (500m^2) is 25 ha. This fact drives the FIRED dataset spikes every 25 ha") +
  ggplot2::facet_wrap(~Year)
ggplot2::ggsave(filename = here::here(figsDir, "fired_welty_areahist_yearfacet.png"), bg = "white")

g <- ggplot2::ggplot(simpleWFDat |> dplyr::filter(SizeCategory == "Small") |> dplyr::filter(Year == 2020)) +
  ggplot2::geom_histogram(aes(x = AreaHA, fill = Dataset), alpha = 0.4, bins = 50, position = "identity") +
  ggplot2::labs(title = "FIRED & Welty datasets fire area histogram: \nFires < 1000 acres (405 ha) in W.US in 2020 (50 bins)",
                caption = "A MODIS pixel (500m^2) is 25 ha. This fact drives the FIRED dataset spikes every 25 ha")
g
d <- ggplot_build(g)$data[[1]]
```
```{r}

sf::st_write(weltyInterestWFSub1000, here::here('data', 'derived', 'welty_jeffries_western_wildfire_sub_1000ac.gpkg'))

```


