Analyze data to get a number on prescribed burns in the Western US & visualize comparisons to low-severity burns
Tyler L. McIntosh
CU Boulder CIRES Earth Lab
Last updated: 1/17/23

This script uses the following naming conventions wherever possible:
 lowerCamelCase for variables
 period.separated for functions
 underscore_separated for files

# Setup

```{r, message = FALSE, warning = FALSE, include = FALSE}
## Libraries ----

#Check the required libraries and download if needed
list.of.packages <- c("tidyverse", #Includes ggplot2, dplyr, tidyr, readr, purrr, tibble, stringr, forcats
                      "terra",
                      "sf",
                      "mapview",
                      "here",
                      "future", "future.apply", "furrr", "doFuture", "progressr", #Futureverse!
                      "tictoc", 
                      "mblm", #Median-based linear models (i.e. thiel-sen)
                      "plyr",
                      "gridExtra", #complex graphics
                      "kableExtra", #nice tables
                      "googlesheets4", #read in google sheet data
                      "scales", #add commas to ggplot axis
                      "tigris", #US data
                      "tmap", #mapping
                      "scales") #add commas to axis

#Install all packages that aren't installed yet
install.packages(setdiff(list.of.packages, rownames(installed.packages())))

#Load all packages
invisible(lapply(list.of.packages, library, character.only = TRUE)) #use 'invisible' to prevent output being displayed

## Clean workspace & set up environment ----
rm(list=ls()) #Ensure empty workspace if running from beginning
here::i_am("code/10_visualize_new_data_conservative.Rmd")
here::here() #Check here location
#OR
#setwd() #Set working directory directly
options(digits = 6) #Set standard decimal print output
options(scipen = 999) #Turn scientific notation on and off (0 = on, 999 = off)

#Ensure necessary directories are created
figsDir <- here::here("figs")
if (!dir.exists(figsDir)){
  dir.create(figsDir)
}
figsDir <- here::here("figs/both_lcms_lcmap")
if (!dir.exists(figsDir)){
  dir.create(figsDir)
}


derivedDatDir <- here::here("data", "derived")
if (!dir.exists(derivedDatDir)){
  dir.create(derivedDatDir)
}

```

# Read in data
Read in the data;
-From Rud
-From rx burn data munging & collating


```{r}
#From script 02
nfporsWithGee <- read_csv(here::here(derivedDatDir, "nfpors_with_gee_lcms_lcmap.csv"))

#NEW data summaries from Rud, includes small fires
yearSummary <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1wUWLin5cXSd9uflna9f_hliz5HEWxOcc7a58mtDziEk/edit#gid=1307905708", sheet = "Year Summary", range = "A1:J12")
stateSummary <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1wUWLin5cXSd9uflna9f_hliz5HEWxOcc7a58mtDziEk/edit#gid=1307905708", sheet = "State Summary", range = "A1:K12")

#OLD data from Rud, events
mtbs <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1cGDeCfF6Gzkf9P4SRHDfqPmZfwyX438jURyzrL49IGk/edit?pli=1#gid=0", sheet = "MTBS Fire Events")
smallFires <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1cGDeCfF6Gzkf9P4SRHDfqPmZfwyX438jURyzrL49IGk/edit?pli=1#gid=0", sheet = "Small Fire Events")

#Filter to just forest rx
nfporsWithGeeForest <- nfporsWithGee |>
  filter(LCMS_PreBurnYear == 1 & LCMAP_PreBurnYear == 4)

#Remove new unburned column
yearSummary <- yearSummary |> dplyr::select(-Unburned)
stateSummary <- stateSummary |> dplyr::select(-Unburned)

```


# Summary stats
```{r}

#Summary stats
totHighGood <- sum(yearSummary$`High Goodfire (CBI High FRG Replacement, HA)`)
totLowGood <- sum(yearSummary$`Low/Mod Goodfire (CBI Low/mod in FRG low/mod, HA)`)
totForFire <- sum(yearSummary$`burnFor (low+mod+high CBI, HA)`)
totRxForest <- sum(dplyr::filter(nfporsWithGeeForest, ACTUALCOMPLETIONYEARNEW < 2021)$TOTALACCOMPLISHMENT_HA)
totAllGood <- sum(totLowGood, totHighGood)

#Summary notes
nfires <- length(unique(mtbs$Event_ID)) + length(unique((smallFires$OBJECTID)))
rxTypes <- unique(nfporsWithGeeForest$TYPENAME)
nNFPORS <- length(unique(nfporsWithGeeForest$objectid__))

summaryStats <- glue::glue("This analysis included {comma(nfires)} total fire event perimeters, and {comma(nNFPORS)} total prescribed fire event instances. \n\n 'Prescribed fire' was defined as including {paste(rxTypes, collapse = ', ')} management actions. \n\n The total area of low/mod goodfire is {comma(round(totLowGood))} ha, while the total area of high goodfire is {comma(round(totHighGood))} ha. These are respectively {round((totLowGood / totForFire) * 100)}% and {round((totHighGood / totForFire) * 100)}% of the total area of forest burned. High good fire is {round((totHighGood / totLowGood) * 100)}% of the amount of low/mod goodfire. \n\n There were {comma(round(totRxForest))} ha of prescribed forest fires. Low/mod goodfire is {round(((totLowGood - totRxForest) / totRxForest) * 100)}% more by area than what was accomplished by prescribed burning. High goodfire is {round((totHighGood / totRxForest) * 100)}% of the area accomplished by prescribed burning. \n\n If we combine high and low/mod good fire, we get {comma(round(totAllGood))} ha of total good fire, which is {round((totAllGood / totForFire) * 100)}% of the total area of forest burned, and {round(((totAllGood - totRxForest) / totRxForest) * 100)}% more by area than what was accomplished by prescribed burning")
print(summaryStats)
sink(here::here(figsDir, "summary.txt"))
cat(glue::glue("Analysis most recent run: {Sys.time()} \n\n\n"))
cat(summaryStats)
sink()


```

# Combine rx & low-CBI from Rud

```{r}

#STATE SUMMARY

stateAll <- nfporsWithGeeForest |>
  dplyr::group_by(STATEABBR) |>
  dplyr::summarise(rxBurnHa = sum(TOTALACCOMPLISHMENT_HA)) |>
  dplyr::left_join(stateSummary, join_by(STATEABBR == STUSPS))


#YEAR SUMMARY

yearAll <- nfporsWithGeeForest |>
  dplyr::group_by(ACTUALCOMPLETIONYEARNEW) |>
  dplyr::summarise(rxBurnHa = sum(TOTALACCOMPLISHMENT_HA)) |>
  dplyr::left_join(yearSummary, join_by(ACTUALCOMPLETIONYEARNEW == year)) |>
  dplyr::rename(year = ACTUALCOMPLETIONYEARNEW) |>
  dplyr::filter(year < 2021) |>
  dplyr::mutate(year = as.character(year))

#A specific version of pivot_longer to do what is specifically required here
pivot.longer <- function(dat) {
  out <- dat |>
    dplyr::rename(goodFire = `Low/Mod Goodfire (CBI Low/mod in FRG low/mod, HA)`,
                  burnFor = `burnFor (low+mod+high CBI, HA)`) %>%
    {if("% low/mod that is goodfire" %in% names(.)) { 
      dplyr::select(., -`% low/mod that is goodfire`, -`% high that is goodfire`, -`% low/mod/high that is goodfire`)
    } else . } %>%
    {if("Low/Mod CBI (HA)" %in% names(.)) { 
      dplyr::select(., -`Low/Mod CBI (HA)`, -`High CBI (HA)`, -`High Goodfire (CBI High FRG Replacement, HA)`)
    } else . } |>
    tidyr::pivot_longer(cols = c(rxBurnHa,
                                 goodFire,
                                 burnFor),
                        names_to =  "datNm",
                        values_to = "hectares") |>
    dplyr::mutate(datNm = factor(datNm, levels = c("rxBurnHa", "goodFire", "burnFor")))
  return(out)
}

#Use function
yearAllLong <- yearAll |> pivot.longer()
stateAllLong <- stateAll |> pivot.longer()
allLong <- yearAll |>
  dplyr::summarise(rxBurnHa = sum(rxBurnHa),
                   `Low/Mod Goodfire (CBI Low/mod in FRG low/mod, HA)` = sum(`Low/Mod Goodfire (CBI Low/mod in FRG low/mod, HA)`),
                   `burnFor (low+mod+high CBI, HA)` = sum(`burnFor (low+mod+high CBI, HA)`)) |>
  pivot.longer() |>
  dplyr::mutate(hectares = round(hectares)) |>
  dplyr::mutate(datNmFull = dplyr::case_when(datNm == 'goodFire' ~ "Good Forest Wildfire",
                                             datNm == 'rxBurnHa' ~ 'Prescribed \nForest Burns',
                                             datNm == 'burnFor' ~ 'Total Forest Wildfire')) |>
  dplyr::mutate(datNm = factor(datNm, levels = c( "burnFor", "goodFire", "rxBurnHa"))) |>
  dplyr::mutate(acres = round(hectares * 2.47105))

```

# Graphics

```{r}

#Set graphic parameters
plotColors <- c("seagreen", "#dc0ab4ff", "gray")
plotLabels <- c("Prescribed Burns", "Good Wildfire", "Total Wildfire")

#Year graphic
yearGraphic <- ggplot2::ggplot(yearAllLong)  +
  geom_col(aes(x = year, y = hectares, fill = datNm),
           position = "dodge", show.legend = FALSE) +
  xlab("Year") +
  ylab("Hectares") +
  labs(title = "Good fire in Western U.S. forests over time (conservative estimate)") +
  scale_y_continuous(labels=comma,
                     sec.axis = sec_axis(trans = ~ . * 2.47105, #add second axis showing acres
                                         name = "Acres",
                                         labels = comma)) +
  theme_light() +
  theme(legend.title = element_blank()) + 
  scale_fill_manual(values = plotColors, labels = plotLabels)
ggplot2::ggsave(plot = yearGraphic, filename = here::here(figsDir, 'cbi_rx_year_fig_v3.png'), units = "in", width = 8, bg = "white")

#State graphic
stateGraphic <- ggplot2::ggplot(stateAllLong)  +
  geom_col(aes(x = STATEABBR, y = hectares, fill = datNm),
            position = "dodge", show.legend = FALSE) +
  xlab("State") +
  ylab("Hectares") +
  labs(title = "Good fire in Western U.S. forests by state (conservative estimate)") +
  scale_y_continuous(labels=comma,
                     sec.axis = sec_axis(trans = ~ . * 2.47105, #add second axis showing acres
                                         name = "Acres",
                                         labels = comma)) +
  theme_light() +
  theme(legend.title = element_blank()) + 
  scale_fill_manual(values = plotColors, labels = plotLabels)
ggplot2::ggsave(plot = stateGraphic, filename = here::here(figsDir, 'cbi_rx_states_fig_v3.png'), units = "in", width = 8, bg = "white")

#Overall graphic
  #Needs different parameters since turned sideways
  plotColors2 <- c("gray", "#dc0ab4ff", "seagreen")
  plotLabels2 <- c("Good Forest Wildfire", "Prescribed Forest Burns", "Total Forest Wildfire")
overallGraphic <- ggplot2::ggplot(allLong)  +
  geom_col(aes(x = datNm, y = hectares, fill = datNm), position = "dodge", show.legend = FALSE) +
  xlab("Year") +
  ylab("Hectares") +
  labs(title = "      2010-2020 Good Fire in Western U.S. Forests (conservative estimate)") +
  scale_y_continuous(labels=comma) +
  theme_void() +
  #theme(plot.title = element_text(hjust = 0.5)) + #make title centered
  scale_fill_manual(values = plotColors2, labels = plotLabels2) +
  geom_text(data = allLong |> filter(datNm != "burnFor"), aes(x = datNm, y = hectares, label = glue::glue("{scales::comma(hectares)} ha\n({scales::comma(acres)} acres)")), hjust = -0.1, colour = "black", size = 4) +
  geom_text(data = allLong |> filter(datNm == "burnFor"), aes(x = datNm, y = hectares, label = glue::glue("{scales::comma(hectares)} ha\n({scales::comma(acres)} acres)")), hjust = 1.1, colour = "white", size = 4) +
  geom_text(aes(x = datNm, y = 0, label = datNmFull), hjust = -0.1, colour = "white", size = 4) +
  coord_flip()
ggplot2::ggsave(plot = overallGraphic, filename = here::here(figsDir, 'overall_v2.png'), units = "in", width = 6, height = 2.5, bg = "white")

```

# Combined graphic

```{r}

#Arranging complex grobs: https://cran.r-project.org/web/packages/gridExtra/vignettes/arrangeGrob.html
lay <- rbind(c(1, NA),
             c(2,3))

combinedPlot <- gridExtra::arrangeGrob(overallGraphic,
                                       stateGraphic + ggplot2::labs(title = NULL),
                                       yearGraphic + ggplot2::labs(title = NULL),
                                       layout_matrix = lay)#,
                                       #top = "Good Fire & Rx in the Western US (conservative estimate)")
ggplot2::ggsave(combinedPlot, filename = here::here(figsDir, 'combined_fig1.png'), units = "in", width = 10)


```

# Tables

```{r}

#Get table with overall numbers
overallTbl <- stateAllLong |> 
  tidyr::pivot_wider(names_from = "datNm",
                     values_from = "hectares") |>
  dplyr::mutate(percGoodOfTot = paste(round(100 * (goodFire / burnFor)), "%", sep = "")) |>
  #dplyr::mutate(percRxOfTot = paste(round(100 * (rxBurnHa / burnFor)), "%", sep = "")) |>
  dplyr::select(-STATEABBR) |>
  dplyr::rename(Group = NAME) |>
  rbind(
    yearAllLong |>
      tidyr::pivot_wider(names_from = "datNm",
                         values_from = "hectares") |>
      dplyr::mutate(percGoodOfTot = paste(round(100 * (goodFire / burnFor)), "%", sep = "")) |>
      dplyr::rename(Group = year)
  ) |>
  rbind(
    allLong |>
      dplyr::select(-datNmFull, -acres) |>
      tidyr::pivot_wider(names_from = "datNm",
                         values_from = "hectares") |>
      dplyr::mutate(percGoodOfTot = paste(round(100 * (goodFire / burnFor)), "%", sep = ""),
                    Group = "All")
  ) |>
  dplyr::mutate(dplyr::across(.cols = rxBurnHa:burnFor, round))

#Create nice table
kable(overallTbl,
      caption = "Good fire in the western U.S. by year, state, and overall (2010-2020) (conservative estimate)",
      format.args = list(big.mark = ","),
      align = rep('l', 5),
      col.names = c("Group","Prescribed burn forest HA", "Good forest wildfire HA", "Total forest wildfire HA", "Good forest wildfire as a percentage of total forest wildfire")) |>
  column_spec(1, width = "2cm") |>
  column_spec(2:5, width = "3cm") |>
  kable_classic(full_width = F, html_font = "Calibri") |>
  save_kable(here::here(figsDir, 'summary_table_v1.png'))

```

# Create CA graphics

```{r california-graphics}


#Pull data by-year for CA, calculate totals
#NEW data summaries from Rud, includes small fires
cbiLowAll <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1wUWLin5cXSd9uflna9f_hliz5HEWxOcc7a58mtDziEk/edit#gid=1307905708", sheet = "CBI lowmod All", range = "A1:N13")
cbiHighAll <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1wUWLin5cXSd9uflna9f_hliz5HEWxOcc7a58mtDziEk/edit#gid=1307905708", sheet = "CBI high all", range = "A1:N13")
cbiGood <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1wUWLin5cXSd9uflna9f_hliz5HEWxOcc7a58mtDziEk/edit#gid=1307905708", sheet = "CBI lowmod FRG lowmod", range = "A1:N13")


#A function to get CA stats in long form from Rud's summaries
get.Ca.Stats <- function(dats, fieldnm) {
  out <- dats |>
  dplyr::select(-`Total (HA)`) |>
  dplyr::filter(NAME == "California") |>
  tidyr::pivot_longer(cols = b2010:b2020,
                      names_to = c("year"),
                      values_to = fieldnm)
  return(out)
}


#Create CA by time dataset
#do rx data
caRxTime <- nfporsWithGeeForest |>
  dplyr::filter(STATEABBR == "CA") |>
  dplyr::group_by(STATEABBR, ACTUALCOMPLETIONYEARNEW) |>
  dplyr::summarise(rxBurnHa = sum(TOTALACCOMPLISHMENT_HA)) |>
  dplyr::rename(year = ACTUALCOMPLETIONYEARNEW) |>
  dplyr::ungroup() |>
  dplyr::select(-STATEABBR)

#Merge to dats from Rud
caTimeStatsLong <- cbiLowAll |>
  get.Ca.Stats("CBILow") |>
  dplyr::left_join(cbiHighAll |> get.Ca.Stats("CBIHigh")) |>
  dplyr::left_join(cbiGood |> get.Ca.Stats("goodFire")) |> #merge all the data for CA
  dplyr::mutate(burnFor = CBILow+CBIHigh, #Get burnFor
                year = as.integer(substr(year, 2, nchar(year)))) |>
  dplyr::mutate(burnFor = burnFor*0.0001,
                goodFire = goodFire*0.0001) |> #convert to HA from m^2
  dplyr::select(-CBILow, -CBIHigh, -NAME, -STUSPS) |>
  dplyr::left_join(caRxTime) |>
  tidyr::pivot_longer(cols = c(rxBurnHa,
                             goodFire,
                             burnFor),
                      names_to =  "datNm",
                      values_to = "hectares") |>
  dplyr::mutate(datNm = factor(datNm, levels = c("rxBurnHa", "goodFire", "burnFor")),
                year = as.character(year))
  


#Get overall CA stats for overall summary graphic
caLong <- stateAllLong |>
  dplyr::filter(STATEABBR == "CA") |>
  dplyr::mutate(datNmFull = dplyr::case_when(datNm == 'goodFire' ~ "Good Forest Wildfire",
                                             datNm == 'rxBurnHa' ~ 'Prescribed \nForest Burns',
                                             datNm == 'burnFor' ~ 'Total Forest Wildfire')) |>
  dplyr::mutate(datNm = factor(datNm, levels = c( "burnFor", "goodFire", "rxBurnHa"))) |>
  dplyr::mutate(acres = round(hectares * 2.47105))


#CA summary graphic
  #Needs different parameters since turned sideways
  plotColors2 <- c("gray", "#dc0ab4ff", "seagreen")
  plotLabels2 <- c("Good Forest Wildfire", "Prescribed Forest Burns", "Total Forest Wildfire")
  
caSummary <- ggplot2::ggplot(caLong) +
  geom_col(aes(x = datNm, y = hectares, fill = datNm), position = "dodge", show.legend = FALSE) +
  xlab("Year") +
  ylab("Hectares") +
  labs(title = "      2010-2020 Good Fire in California Forests (conservative estimate)") +
  scale_y_continuous(labels=comma) +
  theme_void() +
  #theme(plot.title = element_text(hjust = 0.5)) + #make title centered
  scale_fill_manual(values = plotColors2, labels = plotLabels2) +
  geom_text(data = caLong |> filter(datNm != "burnFor"), aes(x = datNm, y = hectares, label = glue::glue("{scales::comma(hectares)} ha\n({scales::comma(acres)} acres)")), hjust = -0.1, colour = "black", size = 4) +
  geom_text(data = caLong |> filter(datNm == "burnFor"), aes(x = datNm, y = hectares, label = glue::glue("{scales::comma(hectares)} ha\n({scales::comma(acres)} acres)")), hjust = 1.1, colour = "white", size = 4) +
  geom_text(aes(x = datNm, y = 0, label = datNmFull), hjust = -0.1, colour = "white", size = 4) +
  coord_flip()  
caSummary

ggplot2::ggsave(plot = caSummary, filename = here::here(figsDir, 'california_summary.png'), units = "in", width = 6, height = 2.5, bg = "white")

#Year graphic
plotColors <- c("seagreen", "#dc0ab4ff", "gray")
plotLabels <- c("Prescribed Burns", "Good Wildfire", "Total Wildfire")
caYearGraphic <- ggplot2::ggplot(caTimeStatsLong)  +
  geom_col(aes(x = year, y = hectares, fill = datNm),
           position = "dodge", show.legend = FALSE) +
  xlab("Year") +
  ylab("Hectares") +
  #labs(title = "Good fire in California forests over time (conservative estimate)") +
  scale_y_continuous(labels=comma,
                     sec.axis = sec_axis(trans = ~ . * 2.47105, #add second axis showing acres
                                         name = "Acres",
                                         labels = comma)) +
  theme_light() +
  theme(legend.title = element_blank()) + 
  scale_fill_manual(values = plotColors, labels = plotLabels)
ggplot2::ggsave(plot = caYearGraphic, filename = here::here(figsDir, 'ca_time.png'), units = "in", width = 6, height = 3, bg = "white")




```