Examine tabular data from GEE to find good case studies
Tyler L. McIntosh
CU Boulder CIRES Earth Lab
Last updated: 12/12/23

This script uses the following naming conventions wherever possible:
 lowerCamelCase for variables
 period.separated for functions
 underscore_separated for files

# Setup

```{r, message = FALSE, warning = FALSE, include = FALSE}
## Libraries ----

#Check the required libraries and download if needed
list.of.packages <- c("tidyverse", #Includes ggplot2, dplyr, tidyr, readr, purrr, tibble, stringr, forcats
                      "terra",
                      "sf",
                      "mapview",
                      "here",
                      "future", "future.apply", "furrr", "doFuture", "progressr", #Futureverse!
                      "tictoc", 
                      "mblm", #Median-based linear models (i.e. thiel-sen)
                      "plyr",
                      "kableExtra", #nice tables
                      "googlesheets4", #read in google sheet data
                      "scales", #add commas to ggplot axis
                      "tigris", #US data
                      "tmap", #mapping
                      "scales") #add commas to axis

#Install all packages that aren't installed yet
install.packages(setdiff(list.of.packages, rownames(installed.packages())))

#Load all packages
invisible(lapply(list.of.packages, library, character.only = TRUE)) #use 'invisible' to prevent output being displayed

## Clean workspace & set up environment ----
rm(list=ls()) #Ensure empty workspace if running from beginning
here::i_am("code/05_explore_case_studies.Rmd")
here::here() #Check here location
#OR
#setwd() #Set working directory directly
options(digits = 6) #Set standard decimal print output
options(scipen = 999) #Turn scientific notation on and off (0 = on, 999 = off)

#Ensure necessary directories are created
figsDir <- here::here("figs")
if (!dir.exists(figsDir)){
  dir.create(figsDir)
}

derivedDatDir <- here::here("data", "derived")
if (!dir.exists(derivedDatDir)){
  dir.create(derivedDatDir)
}

```

# Read in data
Read in the data;
-From Rud: https://docs.google.com/spreadsheets/d/1cGDeCfF6Gzkf9P4SRHDfqPmZfwyX438jURyzrL49IGk/edit?pli=1#gid=0
  (has three tabs of data) - all in HA

```{r}
#From script 03 - all 3 sheets of data
fullDataMTBS <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1cGDeCfF6Gzkf9P4SRHDfqPmZfwyX438jURyzrL49IGk/edit?pli=1#gid=0", sheet = "Full Data")

#Raw MTBS perimeter data
mtbsPerims <- sf::st_read(here::here('data', 'raw', 'mtbs_perimeter_data_1984_2021', 'mtbs_perims_DD.shp'))

```
# Look at case studies

```{r}

#Join the event names by event ID (from raw MTBS polygons) & add in percentages
mtbsNames <- mtbsPerims |>
  sf::st_drop_geometry() |>
  dplyr::select(Event_ID, Incid_Name)
rm(mtbsPerims)

fullDataMTBS <- fullDataMTBS |>
  dplyr::left_join(mtbsNames) |>
  dplyr::mutate(goodPercFor = round(100 * (frcLowCbiLow / preFireFor)),
                goodPercAll = round(100 * (frcLowCbiLow / areaHA)),
                percFor = round(100 * (preFireFor / areaHA)))

#Filter to CA 
caDataMtbs <- fullDataMTBS |>
  dplyr::filter(State == "CA")


# #Get top 10 CA events by size
# caLarge <- caDataMtbs |>
#   dplyr::arrange(plyr::desc(areaHA)) |>
#   head(10)
# 
# #Get top 10 CA events by 'good forest fire' ha
# caGoodHa <- caDataMtbs |>
#   dplyr::arrange(plyr::desc(frcLowCbiLow)) |>
#   head(10)
# 
# 
# #Get top 10 CA events by % 'good fire' in forest and overall
# caGoodPercFor<- caDataMtbs |>
#   dplyr::arrange(plyr::desc(goodPercFor)) |>
#   head(10)
# caGoodPercAll<- caDataMtbs |>
#   dplyr::arrange(plyr::desc(goodPercAll)) |>
#   head(10)


#Get top/bottom N CA data in combined df
#Top N by area, good fire area, percentage good fire of forested area, and percentage good fire of total area
#Also bottom N by percentage of good fire of forested area, and the same but limited to fires over largeLimit ha original forest
topx <- 10
largeLimit <- 10000
caDataMtbs <- caDataMtbs %>%
  dplyr::mutate(haTop = ifelse(Event_ID %in% dplyr::slice_max(., order_by = areaHA, n = topx)$Event_ID, 1, 0),
                goodHaTop = ifelse(Event_ID %in% dplyr::slice_max(., order_by = frcLowCbiLow, n = topx)$Event_ID, 1, 0),
                goodPercForTop = ifelse(Event_ID %in% dplyr::slice_max(., order_by = goodPercFor, n = topx)$Event_ID, 1, 0),
                goodPercForTopLarge = ifelse(Event_ID %in% dplyr::slice_max(dplyr::filter(., preFireFor >= largeLimit), order_by = goodPercFor, n = topx)$Event_ID, 1, 0),
                goodPercAllTop = ifelse(Event_ID %in% dplyr::slice_max(., order_by = goodPercAll, n = topx)$Event_ID, 1, 0),
                goodPercForBottom = ifelse(Event_ID %in% dplyr::slice_min(., order_by = goodPercFor, n = topx)$Event_ID, 1, 0),
                goodPercForBottomLarge = ifelse(Event_ID %in% dplyr::slice_min(dplyr::filter(., preFireFor >= largeLimit), order_by = goodPercFor, n = topx)$Event_ID, 1, 0))

caTop <- caDataMtbs |>
  dplyr::filter(haTop == 1 | goodHaTop == 1 | goodPercForTop == 1 | goodPercAllTop == 1 | goodPercForTopLarge == 1) |>
  dplyr::select(Incid_Name, year, areaHA, preFireFor, percFor, goodPercAll, goodPercFor, haTop, goodHaTop, goodPercForTop, goodPercForTopLarge, goodPercAllTop, goodPercForBottomLarge, goodPercForBottom) |>
  dplyr::arrange(plyr::desc(areaHA))

caBottom <- caDataMtbs |>
  dplyr::filter(haTop == 1 | goodPercForBottomLarge == 1) |>
  dplyr::select(Incid_Name, year, areaHA, preFireFor, percFor, goodPercFor, haTop, goodPercForBottomLarge, goodHaTop) |>
  dplyr::arrange(dplyr::desc(areaHA))

```

# Visualize 'top ten' info

```{r}
#Use kableExtra to create nice tables to view
caTop |>
  `names<-`(c("Incident Name", "Year", "Total ha", "Forest ha", "% forest", "Good fire % of total area", "Good fire % of forest area", glue::glue("In top {topx} by area"), glue::glue("In top {topx} by good fire area"), glue::glue("In top {topx} by good fire % of forest"), glue::glue("In top {topx} by good fire % of forest (fires over {largeLimit}ha only)"), glue::glue("In top {topx} by good fire % of total area"), glue::glue("In bottom {topx} by good fire % of forest (fires over {largeLimit}ha only)"), glue::glue("In bottom {topx} by good fire % of forest"))) |>
  kableExtra::kable(caption = "Most 'good' forest wildfires in CA") |>
  kableExtra::kable_paper(full_width = T, html_font = "Calibri") |>
  kableExtra::column_spec(8, color = ifelse(caTop$haTop == 1, "black", "white")) |>
  kableExtra::column_spec(9, color = ifelse(caTop$goodHaTop == 1, "green", "white")) |>
  kableExtra::column_spec(10, color = ifelse(caTop$goodPercForTop == 1, "green", "white")) |>
  kableExtra::column_spec(11, color = ifelse(caTop$goodPercForTopLarge == 1, "green", "white")) |>
  kableExtra::column_spec(12, color = ifelse(caTop$goodPercAllTop == 1, "green", "white")) |>
  kableExtra::column_spec(13, color = ifelse(caTop$goodPercForBottomLarge == 1, "red", "white")) |>
  kableExtra::column_spec(14, color = ifelse(caTop$goodPercForBottom == 1, "red", "white"))

caBottom |>
  `names<-`(c("Incident Name", "Year", "Total ha", "Forest ha", "% forest", "Good fire % of forest area", glue::glue("In top {topx} by area"), glue::glue("In bottom {topx} by good fire % of forest (fires over {largeLimit}ha only)"), glue::glue("In top {topx} by good fire area"))) |>
  kableExtra::kable(caption = "Most 'non-good' forest wildfires in CA") |>
  kableExtra::kable_paper(full_width = T, html_font = "Calibri") |>
  kableExtra::column_spec(7, color = ifelse(caBottom$haTop == 1, "black", "white")) |>
  kableExtra::column_spec(8, color = ifelse(caBottom$goodPercForBottomLarge == 1, "red", "white")) |>
  kableExtra::column_spec(9, color = ifelse(caBottom$goodHaTop == 1, "green", "white"))

```

# Visualize overall information

```{r}
ggplot2::ggplot(fullDataMTBS) +
  ggplot2::geom_point(aes(x = preFireFor, y = frcLowCbiLow, color = State)) +
  ggplot2::xlab("Pre fire forest") +
  ggplot2::ylab("Good fire: HA low CBI in low/mod regime") +
  ggplot2::labs(title = "Forest vs good fire")
ggplot2::ggplot(fullDataMTBS) +
  ggplot2::geom_histogram(aes(x = goodPercFor)) +
  ggplot2::labs(title = "Good Fire histogram (all)")
ggplot2::ggplot(fullDataMTBS |> dplyr::filter(areaHA > 10000)) +
  ggplot2::geom_histogram(aes(x = goodPercFor)) +
  ggplot2::labs(title = "Good Fire histogram (>10,000 ha)",
                caption = "All fires in western US, 2010-2021") +
  xlab("Percent of forest burned in good fire")
ggplot2::ggplot(fullDataMTBS) +
  ggplot2::geom_histogram(aes(x = goodPercFor)) +
  ggplot2::facet_wrap(~year)
ggplot2::ggplot(fullDataMTBS) +
  ggplot2::geom_point(aes(x = preFireFor, y = goodPercFor, color = State)) +
  ggplot2::xlim(0,1000)


```



